---
title: "Homework 6, Linear Models"
author: "Emil Hafeez (eh2928)"
date: "11/23/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#Load relevant libraries
library(tidyverse)
library(p8105.datasets)
library(modelr)
#Prep for neater output
knitr::opts_chunk$set(
 fig.width = 6,
 fig.asp = 0.8,
 out.width = "80%"
)
#Theming with viridis, minimal
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
#set seed for replicability
set.seed(1)
```

# Problem One

Read in and tidy
```{r read in and tidy, message = F}
homicide_df = 
  read_csv("data/homicide.csv", na = c("", "NA", "Unknown")) %>% 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) %>% 
  filter(
    victim_race %in% c("White", "Black"),
    !city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO")) %>% 
  select(city_state, resolution, victim_age, victim_race, victim_sex)
```

Start with a regression for just one city.

```{r}
baltimore_df =
  homicide_df %>% 
  filter(city_state == "Baltimore, MD")

glm(resolution ~ victim_age + victim_race + victim_sex, 
    data = baltimore_df,
    family = binomial()) %>% 
  broom::tidy() %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(term, OR, starts_with("CI")) %>% 
  knitr::kable(col.names =
                c("Variable", "Odds Ratio Estimate", "Lower CI", "Upper CI"),
              align = "cccc", 
              digits = 3
              )
```

Now, let's run this across multiple cities (all of the 47 remaining).

```{r}
models_results_df = 
  homicide_df %>% 
  nest(data = -city_state) %>% 
  mutate(
    models = 
      map(.x = data, ~glm(resolution ~ victim_age + victim_race + victim_sex, data = .x, family = binomial())),
    results = map(models, broom::tidy)
  ) %>% 
  select(city_state, results) %>% 
  unnest(results) %>% 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) %>% 
  select(city_state, term, OR, starts_with("CI")) 
```

Plot the resultant ORs and CIs of the logistic regression.

```{r}
models_results_df %>% 
  filter(term == "victim_raceWhite") %>% 
  mutate(city_state = fct_reorder(city_state, OR)) %>% 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper)) + 
  geom_hline(yintercept = 1, color = "red") +
  theme(axis.text.x = element_text(angle = 60, vjust = 1, hjust = 1)) +
  scale_y_continuous(breaks = seq(0, 30, 5)) +
  labs(
    title = "Fig. 1: Estimates of the OR for Solving Homicides in US Cities",
    x = "City",
    y = "Odds Ratio Estimate, with CI Bars",
    caption = "Homework 6: Odds Ratio for Solving Homicides where Victim is White (vs reference: Black), Holding Age and Binary Gender Constant")
```

The results show that in many US cities of the 47 for which data are available, the odds of a homicide case ending in arrest are greater when the victim is white compared to Black, holding age of the victim and gender (binary) constant. While there is variability in these estimated which often overlaps the null hypothesis value of OR = 1, most estimates are above 1, and many cities have CIs above 1 which do not overlap the null. This is another modern example of racial injustice in our judicial system, for even if arrest and incarceration is not an effective solution, the apparent racial disparity of whose victims find "justice" as defined by our criminal system feels like yet another example of who is prioritized in our society. 

# Problem Two
```{r}

```


# Problem Three
```{r}

```


